### 2014-08-22 ising/Fibonacci/about.tex

\title{Error Correction in a Phenomenological Fibonacci Anyon Quantum Code}

\author{Simon Burton, Courtney Brell, Steven Flammia}
%\date{\today}

\documentclass[12pt,a4paper]{article}
%\documentclass[11pt, twocolumn]{article}

\usepackage[paper=a4paper,dvips,top=1.5cm,left=1.5cm,right=1.5cm,
    foot=1cm,bottom=1.5cm]{geometry}

%\usepackage{epsf}
\usepackage{amsmath}
\usepackage{color}
\usepackage{natbib}
%\usepackage{cite}

\RequirePackage{amsmath}
\RequirePackage{amssymb}
\RequirePackage{amsthm}
%\RequirePackage{algorithmic}
%\RequirePackage{algorithm}
%\RequirePackage{theorem}
%\RequirePackage{eucal}
\RequirePackage{color}
\RequirePackage{url}
\RequirePackage{mdwlist}

\RequirePackage[all]{xy}
\CompileMatrices
\RequirePackage{hyperref}
\RequirePackage{graphicx}
\RequirePackage[dvips]{geometry}

\begin{document}

\maketitle

\def\Complex {C}
\def\tensor{\otimes}
\def\Tensor{\bigotimes}
\def\bra #1{\langle #1|}
\def\ket #1{|#1\rangle}
\def\braket #1#2{\langle #1|#2 \rangle}

%\def\Set{\widetilde{\text{Set}}}
%\def\Top{\widetilde{\text{Top}}}
%\def\Vec{\widetilde{\text{Vec}}}
%\def\Chain{\widetilde{\text{Chain}}}

%\def\ker{\text{ker}}
%\def\coker{\text{coker}}
%\def\im{\text{im}}

%\def\H{\mathcal{H}}
%\def\H{H}
%\def\S{S}
\def\mathZ{\mathbb{Z}}
\def\mathR{\mathbb{R}}

%\def\nin{\not\in}

%\def\heading #1{\noindent\underline{\Large\bf #1}}
\def\heading #1{\vskip 20pt \noindent\underline{\bf #1}\vskip 5pt}

% ----------------------------------------------------------------------------
%

\heading{Braid group}

Braid group on $n$ strands is the group $B_n$ generated by $n-1$ generators
$\sigma_1, \sigma_2, ... \sigma_{n-1}$ with the following relations:
    $$ \sigma_i \sigma_j = \sigma_j \sigma_i \ \text{when}\ |i-j| > 2, $$
    $$ \sigma_i \sigma_{i+1} \sigma_i =  \sigma_{i+1} \sigma_i \sigma_{i+1}.$$

% ----------------------------------------------------------------------------

\heading{Fibonacci Anyons}

So we have a (linear) representation of the braid group:

    $$ \kappa : B_n \to GL(F_n).$$

Where $F_n$ is the Hilbert state space of $n$ Fibonacci anyons with
trivial total charge.

% ----------------------------------------------------------------------------

\heading{Curve Diagrams}

%Given a topological space $X$ we define the {\it mapping class group} of $X$ as
%the set of homeomorphisms $f:X\to X$ modulo isotopy
%
%    % $$ MCG(X) := \{ f : X \to X \text{s.t.} f \text{is homeomorphism} \} / \sim_{\text{iso}} $$
%    $$ MCG(X) := \{ f : X \to X \} / \sim_{\text{iso}}.$$

Given a topological manifold $M$ we define the {\it mapping class group} of $M$ as
the set of homeomorphisms $f:M\to M$ modulo isotopy

    $$ MCG(M) := \{ f : M \to M \} / \sim_{\text{iso}}.$$

If $M$ has boundary $\partial M$ we require the homeomorphisms to
be the identity on $\partial M$.

In particular we are interested in
 the $n$-punctured disc.
This is a disc $D = \{ x\in \mathR^2 s.t. |x|\leq 1 \} $
with $n$ points removed:

    $$ D_n := D - Q_n,$$

where $Q_n$ is some finite subset of the interior of $D$.

%topologically this is a sphere with $n+1$ holes in it.

It is a theorem that the group $MCG(D_n)$ is isomorphic to $B_n.$
(See \cite{Kassel10}, chapter 1.)

We fix $Q_n$ to be the points $\{(0, i/(n+1)) for i=1,...n\}$
then we define the line $L = \{(0, x) s.t. |x|\leq 1\}.$
Given a representative $f$ of an element of $MCG(D_n)$
we define the
curve diagram as the
image of $L$ under $f$.

It is a theorem (see \cite{Dehornoy02}) that such curve diagrams
correspond uniquely to elements of the mapping class group, and
hence to the braid group $B_n$.

Here we verify the braid relations directly:

\begin{center}
%\includegraphics[width=0.5\textwidth]{mypicture.png}
\includegraphics[width=0.5\textwidth]{curve-braid.eps}
\end{center}

% ----------------------------------------------------------------------------

\heading{Pair-of-pants decompositions}

%We measure charge in regions bounded by curves, ie. elements of
%the first homology group $H_1(D_n).$

%http://arxiv.org/pdf/1002.2816.pdf
%http://www.math.ucsb.edu/~zhenghwa/data/course/lecture_notes/Chap2.pdf
%http://physics.stackexchange.com/questions/93183/about-the-atiyah-segal-axioms-on-topological-quantum-field-theory
%http://mathoverflow.net/questions/359/a-reading-list-for-topological-quantum-field-theory
%http://projecteuclid.org/download/pdf_1/euclid.cmp/1104178138
%http://arxiv.org/pdf/math/9912085v1.pdf

% ----------------------------------------------------------------------------

\bibliography{refs}{}
\bibliographystyle{abbrv}

\end{document}