### 2016-12-06 reresubmit/Response.txt
Some polish on the reply.

We would like to thank the three referees for their comments on our manuscript.  We have previously addressed the substantive concerns of Referees A and B, and in our current submission we have addressed the main concerns of Referee C. We believe that we have now addressed all of the concerns of the three referees and that our manuscript is now appropriate for publication in the Physical Review. We highlight some relevant comments from the referees in support of our manuscript:

Referee A
"The article is very well written and solves a substantial open problem"
"After [revision], the article will be an excellent contribution"

Referee B
"I suggest that the paper is submitted to PRA"

Referee C
"The topic is important both for quantum information science and condensed matter physics"
"I highly suggest the authors to submit the manuscript to PRX"

Response to Referee C:

We thank the referee for their thoughtful comments. We hope that our response clarifies the applicability and scope of our work. Our responses to each comment follow:

(1) It is unclear that the estimated logical error probability
reflects the true logical error probability appropriately. Suppose two
pair creations occur simultaneously nearby, two anyons are braided
with each other, and then syndrome is measured after that. If pair
annihilations are done by the decoder, the original information seems
to be restored according to the authors analysis. But it does not
return to the original one because of the braiding of two anyons. From
what reason, the authors can neglect the braiding process of anyons,
which occurs with a high probability around the threshold? How can the
authors fairly validate their simulation neglecting the braiding
dynamics of the anyons?

Response: In fact, treating the effects that the referee describes was the main challenge of our work. In the model that we used in this paper, the logical state was encoded in the non-trivial topology of the manifold. This meant that only braids that form homologically non-trivial loops will affect the encoded information. In any case, we have simulated completely all the outcomes of braids between anyons (including any effect on the logical state), which was our main technical achievement, and so we have been careful not to neglect any important effects that they may have on the error-correction process.

(2) Do the authors assume that quantum computation is done by anyonic
excitations? If so, we have to consider the dynamics between the
anyons which are employed for quantum computation and undesired
excitations. In the case of the topological codes consisting of qubits
or qudits, we can enlarge the topological defects (and hence the code
distance). However this is not the case for the "physical" anyons
encoding the logical quantum information. If a pair creation occur
nearby it and the excitation is braided around the anyon encoding the
logical quantum information, the logical information would be
destroyed. This seems to worsen the threshold.

Response: Our results apply equally well to a code where the information is stored in the topology of the manifold as to a code where the information is stored in the fusion outcomes of several non-abelian anyons, as would be the case in an anyonic quantum computer. We explicitly treated both cases in our previous work for non-universal non-abelian anyons (Phys. Rev. X 4, 031058). If the information were stored in fusion outcomes of multiple anyons, one might be concerned about braid processes where "noise" anyons braid around the "code" anyons. However, in comparison to topological defects like punctures, such a braid is not a logical error. Since the logical degree of freedom is the joint fusion channel of a pair of anyons, only braid processes that go around the entire pair will contribute to a logical error. A braid that goes around a single code anyon may interact with the charge of that anyon, but it will not interact with the logical state. This is the principal strength of topological fault tolerance for anyonic computation. In our error-correction procedure, we fuse each code anyon with any charges that lie near it, and this reverses any effects of small braids around that code anyon. The fact that one need not enlarge the size of an anyon to increase the distance of the code (as would be the case with punctures) can actually be an advantage in some circumstances by reducing the overhead.

(3) While the introduction of the non-Abelian anyon system is quite
pedagogical and is well mathematically organized, the main
contribution of the manuscript, that is, the proposed efficient
algorithm to sample the non-Abelian anyons on the lattice is described
unclearly. The authors should write down the procedure more explicitly
by using equations and mathematics. The lack of the concrete
description unfortunately make it hard to justify the novelty of the
proposal.

Response: We have addressed this concern by merging the supplementary material, where the main discussion of our algorithm takes place, with the main text, together with a few additional clarifying remarks. We hope that the pseudo code description of our algorithm is a sufficient aid to help non-experts grasp the principles involved. However, we recognize that a complete reimplementation of our work would require a thorough understanding of the prior literature on quantum decoders, and that it might not be completely detailed as a standalone document. Given that other referees (nb Ref A) have commented that the paper was well written, we had to choose between adding additional pedagogical material, or trusting that non-experts would get a coarser but hopefully satisfactory understanding from the current text. Because many of the ideas in our decoder are standard in the literature, we have opted to keep our current description, and we hope this is sufficient.

In summary, we believe we have addressed all of the concerns and that the manuscript is now suitable for publication in Physical Review A.